### 2023-04-04 README.Rmd
update README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
metadata <- rmarkdown::yaml_front_matter("paper.Rmd")
```

# `r metadata$title`

This repo contains the materials to fully reproduce the paper titled "`r metadata$title`" by `r knitr::combine_words(purrr::map_chr(metadata$author, ~.x$name))`.

## Citation

Please cite the paper as:

Li et al. (2023) `r metadata$title`. URL:  https://github.com/TengMCing/lineup_residual_diagnostics